Parallel Information Set Generation for Kriegspiel

### 2012-07-16 sbac/paper.tex
revise intro

\begin{abstract}
Information Set Generation (ISG) is the identification of the set of paths in an
imperfect information game tree that are consistent with a player's
observations.  The ability to reason about the possible game history is
critical to the performance of game-playing agents.  ISG represents a class of
combinatorial search problems which is computationally intensive but difficult
to efficiently parallelize.

@@

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12information}.  Because the operation is
computationally intensive and may be repeatedly applied by a decision-making
agent, its efficiency is a key factor in an agent's overall performance.

The term {\em belief state} is sometimes used interchangeably with information
set to refer to a probability distribution over possible worlds.  The latter
term comes from the game theory community and is preferred here.  A node in a
game tree denotes not only the current state of the game, but also uniquely
defines a path from the initial state or root node.  Thus, a game tree node
implicitly encodes not only the current state of the game but also the complete
history of all decisions made by all players up to that point in the game,
including the outcome of any chance moves such as dice rolls or card shuffling.
Knowing one's information set means knowing all possible game histories.

@@

In the game of Kriegspiel, information set generation is much more complicated.
After the first move by the opponent, a player faces uncertainty about the
location of the opponent's pieces.  Unlike Poker, it is not possible to simply
permute all of the opponent's possible pieces over all of the squares not
occupied by the player's own pieces.  A configuration of pieces for the
opponent is only valid if it can be reached by a legal sequence of moves.  In
the general case, finding the nodes in an information set is a combinatorial
search problem.

In this work, we show that ISG for Kriegspiel can be efficiently parallelized.
The primary contributions of this work are the following. We describe an
approach to efficient parallelization of an application which is not typical of
high performance computing domain. We present scalability results for our
parallel implementation of an information set generation algorithm for the game
of Kriegspiel using problem instances from real games. Further, we demonstrate
the effects of load balancing strategy and grainsize on parallel performance.

The paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
necessary background material, including the rules of Kriegspiel. Next, in
Section~\ref{info} we describe the nature of the information set generation
problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
parallel implementation, including the {\sc Charm++} Parallel State Space
Search Engine (ParSSSE) platform.  We then provide our experimental results in
Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We
conclude with a discussion of future work.

@@

\section{Conclusion and Future Work}
We have shown that information set generation for Kriegspiel can be efficiently
parallelized.  The opportunities for parallelism increase as the problem size
increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes---much larger than any other treatment of the
problem that we are aware of.  Because the algorithm can be efficiently
parallelized, it would be reasonable to explore using it in a full Kriegspiel
player.  It remains to be seen whether these scalability properties hold over a
wide variety of playing styles and problem instances.  We suspect that fruitful
research remains with respect to alternative algorithms that implement variable
ordering heuristics.

>>>>

\begin{abstract}
Information Set Generation (ISG) is the identification of the set of paths in an
imperfect information game tree that are consistent with a player's
observations.  The ability to reason about the possible game history is
critical to the performance of game-playing agents.  ISG represents a class of
combinatorial search problems which is computationally intensive but challenging
to efficiently parallelize.

@@

\section{Introduction}
In imperfect information games, players do not have access to full knowledge of
the world. Examples of imperfect information include hidden cards in
Poker~\cite{billings02challenge} or Bridge~\cite{ginsberg96partition}, hidden
tiles in Scrabble~\cite{richards07opponent}, or hidden pieces in 
Kriegspiel (partially observable chess)~\cite{li94chess}. The game tree nodes
that are indistinguishable to a player because they differ only in the
information that is hidden to the player by rule compose that player's {\em
information set.}  The ability to estimate the value of the possible states and
to reason about the probability distribution over those states is crucial to
playing imperfect information games well~\cite{schofield12hyperplay}.

Information Set Generation (ISG) is an operation that game tree search agents
can use to identify nodes in the game tree that are consistent with their
observations~\cite{richards12information}.  Since the operation is
computationally intensive and may be repeatedly applied by a decision-making
agent, its efficiency is a key factor in an agent's overall performance.

The term {\em belief state} is sometimes used interchangeably with information
set to refer to a probability distribution over possible worlds.  The latter
term comes from the game theory community and is preferred here.  A node in a
game tree denotes not only the current state of the game, but also uniquely
defines a path from the initial state or root node.  Hence, a game tree node
implicitly encodes not only the current state of the game but also the complete
history of all decisions made by all players up to that point in the game,
including the outcome of any chance moves such as dice rolls or card shuffling.
Knowing one's information set means knowing all possible game histories.

@@

However, in the game of Kriegspiel, information set generation is much more complicated.
After the first move by the opponent, a player faces uncertainty about the
location of the opponent's pieces.  Unlike Poker, it is not possible to simply
permute all of the opponent's possible pieces over all of the squares not
occupied by the player's own pieces.  A configuration of pieces for the
opponent is only valid if it can be reached by a legal sequence of moves.  In
the general case, finding the nodes in an information set is a combinatorial
search problem.

In this work, we show that ISG for Kriegspiel can be efficiently parallelized.
The primary contributions of this work are the following:
\begin{itemize}
\item
 We describe an
approach to efficient parallelization of an application which is not typical of
high performance computing domain.
\item
 We present scalability results for our
parallel implementation of an information set generation algorithm for the game
of Kriegspiel using problem instances from real games. We achieve results which are unprecendented in game tree search application. 
\item
We demonstrate
the effects of load balancing strategy and grainsize on the parallel performance of our implementation. 
\end{itemize}

The remainder of the paper is organized as follows.  In Section~\ref{Kriegspiel}, we review
necessary background material, including the rules of Kriegspiel. Next, in
Section~\ref{info}, we describe the nature of the information set generation
problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our
parallel implementation, including the {\sc Charm++} Parallel State Space
Search Engine (ParSSSE) platform.  We then provide our experimental results in
Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We
conclude with a discussion of future work.

@@

\section{Conclusion and Future Work}
We have shown that information set generation for Kriegspiel can be efficiently
parallelized.  The opportunities for parallelism increase as the problem size
increases.  Our algorithm has performed well on problems with information sets
with as many as 65 million nodes --- much larger than any other treatment of the
problem that we are aware of.

Since the algorithm can be efficiently
parallelized, it would be reasonable to explore using it in a full Kriegspiel
player.  It remains to be seen whether these scalability properties hold over a
wide variety of playing styles and problem instances.  We suspect that fruitful
research remains with respect to alternative algorithms that implement variable
ordering heuristics.
